Holomorphic field theories and higher algebra

### 2022-09-22 draft.tex
tiny addition about variant dualites
more minor improvements

The {\em magnetic theory} $\cT_{\sfM}(F, \check{N})$ is the holomorphic QCD with $\check{N}$ colors and $F$ flavors, but with mesons $\check{M}$ where $$V_{\check{M}} = \End(\CC^F) \cong \CC^{\oplus F^2}$$
(and $\check{L}$ for the partner field to $\check{M}$).
We use $\check{A}, \check{B}, \check{Q}, \check{R}, \check{Q}^\dag$, and $\check{R}^\dag$ for the other fields.
The superpotential is $$\check{W} =  \Tr(\check{Q}^\dag \check{M}  \check{Q}),$$
where we compose the linear maps before tracing over~$V_{\check{N}}$.
\end{dfn}

\owen{Please double-check that formula.}
%In this case the auxiliary vector space $Z$ is nontrivial (but transforms trivially under the gauge group as in the definition). 
%Our notation is $q,\Tilde{q}$ for elements of the vector spaces $V, \Bar{V}$ and $m\in Z$.
%The magnetic theory is
%\[
%\cT_{\sfM} (N_c, N_f) \define \cT\big(\SU(N_f - N_c) \, , \, (\CC^{N_f - N_c})^{\oplus N_f} \, , \, \CC^{\oplus N^2_f} \, , \, W = m q \Tilde{q} \big)
%\]
%In the definition of $W = m q \Tilde{q}$ we are using the natural contractions which we can describe explicitly as follows.
%Choose a basis $\{q_a^i\}$ for $V = (\CC^{N_f - N_c})^{\oplus N_f}$, $a = 1,\ldots, N_f - N_c$, $i=1,\ldots N_f$ and a dual basis $\{\Tilde{q}_i^a\}$ for $\Bar{V}$.
%Also, let $\{m_i^j\}$ be a basis for $\CC^{\oplus N_f^2}$. 
%Then the cubic polynomial $W$ is
%\[
%W = \sum_{i,j,a} m^i_j q_i^a \Tilde{q}^i_a .
%\]

The usual Seiberg duality states how certain ``electric'' and ``magnetic'' versions of SQCD should be equivalent.
The electric and magnetic holomorphic QCDs are precisely the holomorphic twists of those that appear in Seiberg duality,
and so Richard Eager suggested that Seiberg duality should survive after taking the holomorphic twist.

@@

This claim should seem rather surprising.
In many ways these theories look, on their face, like very different theories:
the gauge groups $SL(N)$ and $SL(\check{N})$ are (usually) different, and
one side has mesons and the other does not.
But this feature also makes the conjecture powerful.

The statement should be interpreted as an equivalence of factorization algebras valued in $\ZZ/2$-graded cochain complexes since the field theories are only $\ZZ/2$-graded.

@@

As we discuss in a moment, holomorphic QCD can be quantized using the techniques of \cite{BW work}.
There are then two further pieces of evidence when these quantum theories are examined:
\begin{enumerate}
\item[(2)] There are {\em flavor} symmetries that act on each theory, and these can be matched under the duality. Under quantization, these have isomorphic anomalies.
\item[(3)] There are {\em spacetime} symmetries that act on each theory, notably isometries, and on the classical holomorphic theories, these enhance to an action of holomorphic vector fields. Under quantization, these have isomorphic anomalies.
\end{enumerate}
This triple of evidence is analogous (and sometimes stronger) than results for Seiberg duality for SQCD. 
We now turn to fleshing out these assertions.

@@

\owen{I want to describe here, at least briefly, what the moduli stack is. I'm not sure the stacks are equivalent, but I think they have equivalent affinizations. I wonder if we can exhibit an equivalence of the "quantum" affinizations.}

@@

In studying questions about the classical vacua of supersymmetric field theories, however, physicists have not typically worked with the {\em stack}.
Instead, they have worked with a GIT quotient.
For instance, in the setting of a gauge theory with moduli stack of classical vacua $V/G$, it means they work with
\[
V\Sslash G = \Spec(\cO(V)^G),
\]
the scheme whose ring of functions are the $G$-invariant functions on the representation~$V$.
This affine scheme is known as the {\em affinization} of the stack $V/G$, as it is the best affine approximation to the stack (i.e., apply the left adjoint to the inclusion of affine schemes among stacks).
In practice physicists work ordinary schemes (i.e., without any derived directions), although one can also study the derived affine scheme that is the best approximation to a derived stack.
We will call this GIT quotient the moduli {\em space} of classical vacua and denote it by~$M_{\rm vac}(\cT)$.
(Note that the font of M changes between stack and space.)

@@

\begin{rmk}
The physicists' focus on the moduli space is not unreasonable:
one must make measurements --- i.e., evaluate functions --- to identify the state of a physical system.
The moduli space captures what can be seen by measurements (notably here, the ``vacuum expectation values'' of different vacua).
On the other hand, the moduli stack knows each orbit, which provides a distinct solution to the equations of motion, whereas the space cannot distinguish between some vacua.
Moreover, the stack encodes much more data about the theory near each vacuum.
For instance, it remembers ``enhanced symmetries'' (such as the residual gauge fields) at each vacuum.
\end{rmk}

For the holomorphic QCD theory $\cT(F,N)$ without mesons, 
the derived stack of classical vacua is 
\[
\cM_{\rm vac}(F,N) \cong T^*[1] \left(V_Q \oplus V_Q^*/SL(N)\right).
\]
By contrast the moduli space $M_{\rm vac}$ is a variety already well-studied in classical algebraic geometry.
Before we jump to the full space, let's examine a simpler case involving just the quarks.
A vacuum for the quark field is an element of $V_Q = \Hom(\CC^F, \CC^N)$,
and it is a classic result of invariant theory that
\[
\Hom(\CC^F, \CC^N)\Sslash SL(N) \cong \widehat{{\rm Gr}(F, N)} \subset \Lambda^N \CC^F
\]
where ${\rm Gr}(F, N)$ denotes the Grassmannian of $N$-planes in $\CC^F$ and the hat indicates we take the affine cone with respect to this standard Pl\"ucker embedding.
For the full problem, we have both quarks and antiquarks, so
\begin{align*}
M_{\rm vac}(F,N) 
&\cong V_Q \oplus V_Q^*\Sslash SL(N) \\
&= \Hom(\CC^F, \CC^N) \oplus \Hom(\CC^N, \CC^F)\Sslash SL(N),
\end{align*}
whose ring of functions is also computed by the fundamental theorems of invariant theory.
Those theorems tell us that the GIT quotient is a closed conical subvariety
\[
M_{\rm vac}(F,N)  \hookrightarrow \Lambda^N \CC^F \times \Lambda^N (\CC^F)^* \times \End(\CC^F)
\]
where a pair of matrices $(X,Y) \in \Hom(\CC^F, \CC^N) \times \Hom(\CC^N, \CC^F)$ goes to $(\Lambda^N X, \Lambda^N Y, YX)$, with $\Lambda^N X$ denoting the wedge product of the rows of $X$ (and of columns for $Y$).
The image thus satisfies both sets of Pl\"ucker relations, the condition that the composite $YX$ has rank at most $N$, and some further relations, given by the Second Fundamental Theorem.
In other words,
\[
M_{\rm vac}(F,N) \hookrightarrow \widehat{{\rm Gr}(F, N)} \times \widehat{{\rm Gr}(F, N)} \times {\rm R}(F,N)
\]
with ${\rm R}(F,N) \subset \End(\CC^F)$ the matrices with rank bounded above by~$N$.
The moduli space for the magnetic theory is a bit more complicated but follows from invariant theory as well.

@@

This relationship, which holds already for the supersymmetric theory before twisting, was one of the original pieces of evidence offered by Seiberg for his duality.
One disocvers the need for mesons and the superpotential in order to obtain this result.

>>>>

The {\em magnetic theory} $\cT_{\sfM}(F, \check{N})$ is the holomorphic QCD with $\check{N}$ colors and $F$ flavors, but with mesons $\check{M}$ where $$V_{\check{M}} = \End(\CC^F) \cong \CC^{\oplus F^2}$$
(and $\check{L}$ for the partner field to $\check{M}$).
We use $\check{A}, \check{B}, \check{Q}, \check{R}, \check{Q}^\dag$, and $\check{R}^\dag$ for the other fields.
The superpotential is $$\check{W} =  \Tr(\check{Q}^\dag \check{M}  \check{Q}),$$
where we compose the linear maps before tracing over~$V_{\check{N}}$.
\owen{Please double-check that formula.}
\end{dfn}

Note that the lower bound on $N$ and $\check{N}$ is simply to ensure that we have nontrivial gauge groups.
%In this case the auxiliary vector space $Z$ is nontrivial (but transforms trivially under the gauge group as in the definition). 
%Our notation is $q,\Tilde{q}$ for elements of the vector spaces $V, \Bar{V}$ and $m\in Z$.
%The magnetic theory is
%\[
%\cT_{\sfM} (N_c, N_f) \define \cT\big(\SU(N_f - N_c) \, , \, (\CC^{N_f - N_c})^{\oplus N_f} \, , \, \CC^{\oplus N^2_f} \, , \, W = m q \Tilde{q} \big)
%\]
%In the definition of $W = m q \Tilde{q}$ we are using the natural contractions which we can describe explicitly as follows.
%Choose a basis $\{q_a^i\}$ for $V = (\CC^{N_f - N_c})^{\oplus N_f}$, $a = 1,\ldots, N_f - N_c$, $i=1,\ldots N_f$ and a dual basis $\{\Tilde{q}_i^a\}$ for $\Bar{V}$.
%Also, let $\{m_i^j\}$ be a basis for $\CC^{\oplus N_f^2}$. 
%Then the cubic polynomial $W$ is
%\[
%W = \sum_{i,j,a} m^i_j q_i^a \Tilde{q}^i_a .
%\]

Seiberg duality states how certain ``electric'' and ``magnetic'' versions of SQCD should be equivalent.
The electric and magnetic holomorphic QCDs are precisely the holomorphic twists of those that appear in Seiberg duality,
and so Richard Eager suggested that Seiberg duality should survive after taking the holomorphic twist.

@@

\begin{rmk}
The statement is an equivalence of factorization algebras valued in $\ZZ/2$-graded cochain complexes since the field theories are only $\ZZ/2$-graded.
\end{rmk}

This claim should seem rather surprising.
In many ways these theories look, on their face, like very different theories:
the gauge groups $SL(N)$ and $SL(\check{N})$ are (usually) different, and
one side has mesons and the other does not.
Of course, relating quite different theories also makes the conjecture powerful.

@@

As we discuss in a moment, holomorphic QCD can be quantized using the techniques of \cite{BW work}.
There are then two further pieces of evidence when these quantum theories are examined:
\begin{enumerate}
\item[(2)] There are {\em flavor} symmetries by $SL(F)$ that act on the quarks and antiquarks of each theory, 
and these can be matched under the duality. 
For the classical holomorphic theories, these symmetries enhance to an action of holomorphic $\sl(F)$-valued functions on the quark and antiquark fields.
Under quantization, these have isomorphic anomalies.
\item[(3)] There are {\em spacetime} symmetries that act on each theory, notably isometries, 
and on the classical holomorphic theories, these enhance to an action of holomorphic vector fields. Under quantization, these have isomorphic anomalies.
\end{enumerate}
This triple of evidence is analogous (and sometimes stronger) than results for Seiberg duality for SQCD. 
We now turn to fleshing out these assertions.

@@

@@

In studying questions about the classical vacua of supersymmetric field theories, however, physicists have not typically worked with the {\em stack}.
Instead, they have worked with a GIT quotient.
For instance, if the moduli stack of classical vacua were $V/G$, it means physicists work with
\[
V\Sslash G = \Spec(\cO(V)^G),
\]
the scheme whose ring of functions are the $G$-invariant functions on the representation~$V$.
This affine scheme is known as the {\em affinization} of the stack $V/G$, as it is the best affine approximation to the stack (i.e., apply the left adjoint to the inclusion of affine schemes among stacks).
We will call this GIT quotient the moduli {\em space} of classical vacua and denote it by~$M_{\rm vac}(\cT)$.
(Note that the font of M changes between stack and space.)

\begin{rmk}
In practice physicists work ordinary schemes (i.e., without any derived directions), 
although one can also study the derived affine scheme that is the best approximation to a derived stack.
Below we will always provide an ordinary scheme as the moduli {\em space}.
\end{rmk}

@@

\begin{rmk}
The physicists' focus on the moduli space is not unreasonable:
one must make measurements --- i.e., evaluate functions --- to identify the state of a physical system.
The moduli space captures what can be seen by measurements (notably here, the ``vacuum expectation values'' of different vacua).
In other words, the possible classical states --- viewed as suitable linear functionals on the algebra of observables --- are parametrized by the underlying affine scheme of the derived moduli stack.
On the other hand, the moduli stack knows each orbit, which provides a distinct solution to the equations of motion,
and hence records much more about the classical theory; 
the space cannot distinguish between some vacua because it collapses orbits.
Moreover, the stack encodes much more data about the theory near each vacuum.
For instance, it remembers ``enhanced symmetries'' (such as the residual gauge fields) at each vacuum.
\end{rmk}

For the holomorphic QCD theory $\cT(F,N)$ without mesons, 
the derived stack of classical vacua is 
\[
\cM_{\rm vac}(F,N) \cong T^*[1] \left(V_Q \oplus V_Q^*/SL(N)\right).
\]
By contrast the moduli space $M_{\rm vac}$ is a variety already well-studied in classical algebraic geometry.
Before we jump to the full space, let's examine a simpler case involving just the quarks.
A vacuum for the quark field is an element of $V_Q = \Hom(\CC^F, \CC^N)$,
and it is a classic result of invariant theory that
\[
\Hom(\CC^F, \CC^N)\Sslash SL(N) \cong \widehat{{\rm Gr}(F, N)} \subset \Lambda^N \CC^F
\]
where ${\rm Gr}(F, N)$ denotes the Grassmannian of $N$-planes in $\CC^F$ and the hat indicates we take the affine cone with respect to this standard Pl\"ucker embedding.
For the full problem, we have both quarks and antiquarks, so
\begin{align*}
M_{\rm vac}(F,N) 
&\cong V_Q \oplus V_Q^*\Sslash SL(N) \\
&= \Hom(\CC^F, \CC^N) \oplus \Hom(\CC^N, \CC^F)\Sslash SL(N),
\end{align*}
whose ring of functions is also computed by the fundamental theorems of invariant theory.
Those theorems tell us that the GIT quotient is a closed conical subvariety
\[
M_{\rm vac}(F,N)  \hookrightarrow \Lambda^N \CC^F \times \Lambda^N (\CC^F)^* \times \End(\CC^F)
\]
where a pair of matrices $(X,Y) \in \Hom(\CC^F, \CC^N) \times \Hom(\CC^N, \CC^F)$ goes to $(\Lambda^N X, \Lambda^N Y, YX)$, with $\Lambda^N X$ denoting the wedge product of the rows of $X$ (and of columns for $Y$).
The image thus satisfies both sets of Pl\"ucker relations, the condition that the composite $YX$ has rank at most $N$, and some further relations, given by the Second Fundamental Theorem.
\owen{I'm trying to find a completely careful and explicit statement of the $SL(N)$ case.}
In other words,
\[
M_{\rm vac}(F,N) \hookrightarrow \widehat{{\rm Gr}(F, N)} \times \widehat{{\rm Gr}(F, N)} \times {\rm R}(F,N)
\]
with ${\rm R}(F,N) \subset \End(\CC^F)$ the matrices with rank bounded above by~$N$.
The moduli space for the magnetic theory is a bit more complicated but follows from invariant theory as well.

@@

This relationship, which holds already for the supersymmetric theory before twisting, was one of the original pieces of evidence offered by Seiberg for his duality.
The proof boils down to checking that the mesons and the superpotential on the magnetic side ensure the isomorphism.
